---
title: "Page count summary"
author: "`r author`"
date: "`r Sys.Date()`"
output: markdown_document
---


## ESTC/ECCO comparison

```{r ecco, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}
# Compare page count info between ESTC and ECCO
df <- df.preprocessed
df.orig <- readRDS("data/unified/polished/df.raw.Rds")

# Read ECCO dump
ecco <- read_ecco(version = 2)

# Polish doc ID
df$id <- df$system_control_number
df$id <- gsub("\\(CU-RivES\\)", "", df$id)

# Add ECCO page counts to our data
df$pagecount.ecco <- ecco[match(df$id, as.character(ecco$id)), "totalPages"]
df$ecco.id <- ecco[match(df$id, as.character(ecco$id)), "id"]
# Add info on whether the page count was estimated or not
df$pagecount.estimated <- !is.na(df$pagecount) & is.na(df$pagecount.orig)

# Augmented pagecounts 
o <- subset(df, pagecount.estimated)$original_row
comptab <- cbind(
	estc.id = as.character(df[match(o, df$original_row),"system_control_number"]),
	ecco.id = as.character(df[match(o, df$original_row), "ecco.id"]),	
	gatherings = as.character(df[match(o, df$original_row),"gatherings"]),
	physical_extent = df.orig[match(o, df.orig$original_row),"physical_extent"],	
	estc = df[match(o, df$original_row),"pagecount"],
	ecco = df[match(o, df$original_row),"pagecount.ecco"],
	estc.volcount = df[match(o, df$original_row),"volcount"],
	estc.volnumber = df[match(o, df$original_row),"volnumber"],
	ecco.hits = rep(NA, length(o)),
	year = df[match(o, df$original_row),"publication_year"],	
	author = df[match(o, df$original_row),"author"],	
	title = as.character(df[match(o, df$original_row),"title"])
	)
# Exclude cases with no ECCO ID	
comptab[which(comptab[, "ecco.id"] == "NA"), "ecco.id"] <- NA
comptab <- comptab[!is.na(comptab[, "ecco.id"]),]	
# Sum up ECCO hits
for (i in 1:nrow(comptab)) {
  myid <- comptab[i,"ecco.id"]
  dfs <- subset(df, id == myid)
  ecco.set <- subset(ecco, id == myid)  
  if (is.na(dfs$volcount) & !is.na(dfs$volnumber)) {
    # Pick the ecco pages just for this particular volume (volcount NA; volnumber given)
    # assuming that the ECCO hits are in same order as the real volumes
    # (this can be assumed to be a good approximation anyway)
    comptab[i, "ecco"] <- sum(ecco.set$totalPages[dfs$volnumber], na.rm = TRUE)
  } else {
    comptab[i, "ecco"] <- sum(ecco.set$totalPages, na.rm = TRUE)
  }
  comptab[i, "ecco.hits"] <- nrow(ecco.set)
}

write.table(comptab, file = "~/tmp/estc_ecco_augmented.csv", quote = F, row.names = F, sep = "\t")

# -------------------------------------------------------------------

# The differing pagecounts in ESTC/ECCO
df2 <- df
df2 <- subset(df, !pagecount.estimated)


# This takes long time to run, so only do once
skip <- FALSE
# TODO: should be streamlined, this whole Rmd doc
if (!skip) {
  comptab <- cbind(
	estc.id = as.character(df2[,"system_control_number"]),
	ecco.id = as.character(df2[, "ecco.id"]),	
	gatherings = df2[,"gatherings"],
	physical_extent = df.orig[match(df2$original_row, df.orig$original_row), "physical_extent"],	
	estc = df2[,"pagecount"],
	ecco = df2[,"pagecount.ecco"],
	estc.volcount = df2[,"volcount"],
	estc.volnumber = df2[,"volnumber"],
	ecco.hits = rep(NA, nrow(df2)),	
	year = df2[,"publication_year"],	
	author = df2[,"author"],	
	title = df2[,"title"]	
	)
  # Exclude cases with no ECCO ID	
  comptab[which(comptab[, "ecco.id"] == "NA"), "ecco.id"] <- NA
  comptab <- comptab[!is.na(comptab[, "ecco.id"]),]

  # Sum up all ECCO hits
  for (i in 1:nrow(comptab)) {
    myid <- comptab[i,"ecco.id"]
    dfs <- subset(df2, id == myid)
    if (is.na(dfs$volcount) & !is.na(dfs$volnumber)) {
      # Pick the ecco pages just for this particular volume (volcount NA; volnumber given)
      # assuming that the ECCO hits are in same order as the real volumes
      # (this can be assumed to be a good approximation anyway)
      ecco.set <- subset(ecco, id == myid)
      # If ECCO lists multiple volumes then try to pick the indicated volume
      comptab[i, "ecco"] <- sum(ecco.set$totalPages[dfs$volnumber], na.rm = TRUE)
      comptab[i, "ecco.hits"] <- nrow(ecco.set)

    } else {
      comptab[i, "ecco"] <- sum(subset(ecco, id == myid)$totalPages, na.rm = TRUE)
    }
  }
  comptab <- comptab[!comptab[, "estc"] == comptab[, "ecco"],]
  o <- order(abs(comptab[, "estc"] - comptab[, "ecco"]), decreasing = TRUE)
  comptab <- comptab[o,]

  write.table(comptab, file = "~/tmp/estc_ecco_pagecount_mismatches.csv",
		     quote = F, row.names = F, sep = "\t")

  saveRDS(comptab, file = "comptab.Rds")
  
} else {

  comptab <- readRDS("comptab.Rds")

}

#------------------------------------------------------------------------

# Percentage of total ESTC pages vs. ECCO pages
# across the shared docs
coms <- intersect(ecco$id, df$id)
total.match <- sum(subset(df, id %in% coms)$pagecount, na.rm = TRUE) / sum(subset(ecco, id %in% coms)$totalPages, na.rm = TRUE)

# Prepare for the figure below
dfs <- subset(df, !is.na(pagecount.ecco) & !is.na(pagecount))
```

There are `r nrow(ecco)` ECCO documents and `r sum(!ecco$id == "", na.rm = TRUE)` have an ESTCID. Some of the ECCO entries refer to the same ESTCID; `r sum(as.character(ecco$id) %in% df$id)` ECCO documents have a direct match in ESTC. The number of unique ECCO documents that have ESTC ID  is `r length(unique(as.character(ecco$id)))` and `r round(100*mean(unique(as.character(ecco$id)) %in% df$id), 1)`% of these have a match in ESTC.

Total page count for the matched ESTC documents is `r round(100 * total.match, 1)`% of the total page count over the same ECCO documents. This quantifies the correspondence between ESTC and ECCO.

Comparison between the page counts available in ESTC and ECCO helps to quantify the accuracy our automated page count cleaning and estimation procedure. Some ESTC page counts are missing in the original data and have been augmented based on predefined estimates for [single volume](mean_pagecounts_singlevol.csv), [multi-volume](mean_pagecounts_multivol.csv) and [issues](mean_pagecounts_issue.csv), calculated from those documents where original page count info is available. In the ESTC/ECCO comparison, where pagecounts are available for both data sets, `r round(100 * mean(dfs$pagecount.estimated),2)`% of the pagecounts are based on estimates; pagecount for `r sum(df$pagecount.from.ecco, na.rm = TRUE)` ESTC documents has been replaced or augmented from ECCO, improving the correlation.


```{r ecco2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=9, out.width="500px"}
library(ggplot2)
theme_set(theme_bw(20))
p <- ggplot(dfs, aes(x = pagecount, y = pagecount.ecco)) +
       geom_point(aes(color = pagecount.estimated)) +
       scale_x_log10() +
       scale_y_log10() +       
       xlab("Page count ESTC") +
       ylab("Page count ECCO") +
       ggtitle(paste("ECCO/ESTC page count comparison (n = ", nrow(dfs), ")", sep = "")) +
       geom_abline(intercept = 0, slope = 1)
print(p)
```



