---
title: "Page count summary"
author: "`r author`"
date: "`r Sys.Date()`"
output: markdown_document
---


## ESTC/ECCO comparison

```{r ecco, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}
# Compare page count info between ESTC and ECCO
df <- df.preprocessed

# Read ECCO dump
ecco <- read.csv("ecco.csv.gz")

# Polish doc ID
df$id <- df.preprocessed$system_control_number
df$id <- gsub("\\(CU-RivES\\)", "", df$id)

# Add ECCO page counts to our data
df$pagecount.ecco <- ecco[match(df$id, as.character(ecco$ESTCID)), "totalPages"]
# Add info on whether the page count was estimated or not
df$pagecount.estimated <- !is.na(df$pagecount) & is.na(df$pagecount.orig)
```

There are `r sum(as.character(ecco$ESTCID) %in% df$id)` ECCO documents (`r round(100*mean(as.character(ecco$ESTCID) %in% df$id), 1)`%) with a direct match in ESTC in our data based on the system control number.

Comparison between the page counts available in ESTC and ECCO can be used to validate our automated page count cleaning and estimation procedure.

```{r ecco2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=9, out.width="500px"}
library(ggplot2)
theme_set(theme_bw(20))
dfs <- subset(df, !is.na(pagecount.ecco) & !is.na(pagecount))
p <- ggplot(dfs, aes(x = pagecount, y = pagecount.ecco)) +
       geom_point(aes(color = pagecount.estimated)) +
       scale_x_log10() +
       scale_y_log10() +       
       xlab("Page count ESTC") +
       ylab("Page count ECCO") +
       ggtitle(paste("ECCO/ESTC page count comparison (n = ", nrow(dfs), ")", sep = "")) +
       geom_abline(intercept = 0, slope = 1)
print(p)
```


## Page counts

Page count: distribution of document sizes

```{r pagecountstat, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}
df <- df.preprocessed

  theme_set(theme_bw(15))

  # Single-volume docs
  dff <- filter(df, volcount == 1 & is.na(volnumber))

  dff2 <- dff %>% group_by(gatherings, 
		    	     pagecount) %>%
			     tally()

  dff3 <- dff %>% group_by(gatherings) %>%
		    summarize(mean = mean(pagecount, na.rm = T), 
		              median = median(pagecount, na.rm = T))

  p <- ggplot(dff2, aes(y = gatherings, x = pagecount)) 
  p <- p + geom_point(aes(size = n))
  p <- p + geom_point(data = dff3, aes(y = gatherings, x = mean), col = "red", size = 3)
  p <- p + geom_point(data = dff3, aes(y = gatherings, x = median), col = "blue", size = 3)
  p <- p + scale_x_log10(breaks = c(1, 10, 100, 1000))
  p <- p + xlab("Total page count (blue: median; red: mean)")
  p <- p + ylab("Document size")
  p <- p + ggtitle(paste("Pages: single-volume documents (n=", nrow(dff), ")", sep = ""))
  p1 <- p 

  # Multi-volume docs
  theme_set(theme_bw(15))
  dff <- filter(df, 
 	   (volcount > 1 | 
	   #(items == 1 & !is.na(volnumber))) # include when items info is again available
	   (!is.na(volnumber)))	   
	   #pagecount > 10
	   )
  dff2 <- dff %>% group_by(gatherings, 
		    	     pagecount) %>%
		    	     tally()

  dff3 <- dff %>% group_by(gatherings) %>%
		    summarize(mean = mean(pagecount, na.rm = T), 
		     median = median(pagecount, na.rm = T))

  p <- ggplot(dff2, aes(y = gatherings, x = pagecount)) 
  p <- p + geom_point(aes(size = n))
  p <- p + geom_point(data = dff3, aes(y = gatherings, x = mean), col = "red", size = 3)
  p <- p + geom_point(data = dff3, aes(y = gatherings, x = median), col = "blue", size = 3)
  p <- p + scale_x_log10(breaks = c(1, 10, 100, 1000))
  p <- p + xlab("Total page count (blue: median; red: mean)")
  p <- p + ylab("Document size")
  p <- p + ggtitle(paste("Pages: multi-volume documents (n=", nrow(dff), ")", sep = ""))
  p2 <- p 

library(gridExtra)
grid.arrange(p1, p2, nrow = 1)
```


## Documents with missing pages over years 

```{r missingpages, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=4}
library(dplyr)
df2 <- df %>% group_by(publication_decade, gatherings) %>% summarize(na = sum(is.na(pagecount.orig) | (!pagecount.orig == pagecount)))
df2$na[df2$na == 0] <- NA
library(ggplot2)
theme_set(theme_bw(15))
p <- ggplot(df2, aes(x = publication_decade, y = gatherings, size = na))
p <- p + geom_point()
p <- p + xlim(c(min(na.omit(df$publication_decade)), 1800))
p <- p + ggtitle(paste("Documents with missing page counts (original; n=", sum(na.omit(df2$na)), ")", sep = ""))
p <- p + xlab("Publication decade")
p <- p + ylab("Document dimension (estimated)")
print(p)

library(dplyr)
df2b <- df %>% group_by(publication_decade, gatherings) %>% summarize(na = sum(is.na(pagecount)))
df2b$na[df2b$na == 0] <- NA
library(ggplot2)
theme_set(theme_bw(15))
p <- ggplot(df2b, aes(x = publication_decade, y = gatherings, size = na))
p <- p + geom_point()
p <- p + xlim(c(min(na.omit(df$publication_decade)), 1800))
p <- p + ggtitle(paste("Documents with missing page counts (after adding averages; n=", sum(na.omit(df2b$na)), ")", sep = ""))
p <- p + xlab("Publication decade")
p <- p + ylab("Document dimension (estimated)")
print(p)
```


## Estimated paper consumption

Note: there are `r length(which(is.na(df$cm2) & (!is.na(df$gatherings.original) | !is.na(df$width.original) | !is.na(df$height.original))))` documents that have some dimensions information but sheet area information could not be calculated for some reason. 

```{r paperconsumption, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=6}
library(dplyr)
library(ggplot2)
df <- df.preprocessed
df2 <- df %>% group_by(publication_year) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
library(sorvi)
p <- regression_plot(paper ~ publication_year, df2) 
p <- p + ggtitle("Total annual paper consumption")
p <- p + xlab("Year")
p <- p + ylab("Paper consumption")
print(p)

# Year vs paper with text
#p <- ggplot(df2, aes(y = paper, x = publication_year))
#p <- p + geom_text(aes(label = publication_year))
#p <- p + ggtitle("Total annual paper consumption")
#p <- p + xlab("Year")
#p <- p + ylab("Paper consumption")
#print(p)

p <- regression_plot(n ~ publication_year, df2) 
p <- p + ggtitle("Total annual documents")
p <- p + xlab("Year")
p <- p + ylab("Documents (n)")
print(p)
```

```{r paperconsumption2b, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=5}
df <- df.preprocessed
df2 <- df %>% group_by(publication_year, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
df2 <- filter(df2, gatherings %in% names(which(table(df2$gatherings) >= 50)))
p <- ggplot(df2, aes(y = paper, x = publication_year, group = gatherings, color = gatherings))
p <- p + geom_point()
#p <- p + geom_line()
#p <- p + scale_y_log10()
p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Annual paper consumption by gatherings")
p <- p + xlab("Year")
p <- p + ylab("Paper consumption")
p <- p + scale_color_discrete(guide = guide_legend(title = "Doc. Size"))
print(p)

p <- ggplot(df2, aes(y = n, x = publication_year, group = gatherings, color = gatherings))
p <- p + geom_point()
#p <- p + geom_line()
#p <- p + scale_y_log10()
p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Annual title count by gatherings")
p <- p + xlab("Year")
p <- p + ylab("Documents (n)")
print(p)




```{r pagecounts-gatherings-relab, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=5}
df <- df.preprocessed
df2 <- df %>% group_by(publication_year, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
df2 <- filter(df2, gatherings %in% names(which(table(df2$gatherings) >= 50)))

p <- ggplot(df2, aes(x = publication_year, fill = gatherings)) + geom_bar(position = "stack", stat = "bin")
p <- p + ggtitle("Gatherings counts over time")
print(p)

p <- ggplot(df2, aes(x = publication_year, fill = gatherings)) + geom_bar(position = "fill")
p <- p + ggtitle("Gatherings proportions over time")
print(p)    

p <- ggplot(df2, aes(y = n, x = publication_year, group = gatherings, color = gatherings))
p <- p + geom_point()
#p <- p + geom_line()
#p <- p + scale_y_log10()
p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Annual title count by gatherings")
p <- p + xlab("Year")
p <- p + ylab("Documents (n)")
print(p)
```

```{r paperconsumption2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=13, fig.height=6}
df <- df.preprocessed
df2 <- df %>% group_by(publication_year) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
p <- ggplot(df2, aes(x = publication_year, y = paper/n))
p <- p + geom_point()
p <- p + geom_line()
p <- p + ggtitle("Average paper consumption per document")
p <- p + xlab("Year")
p <- p + ylab("Average paper consumption per document")
print(p)
```



## Pamphlets vs. Books

```{r doctypes, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=13, fig.height=6}
library(ggplot2)
library(dplyr)
df <- df.preprocessed
df$type <- rep(NA, nrow(df))
df$type[df$pagecount > 32] <- "book"
df$type[df$pagecount <= 32] <- "pamphlet"
df$type <- factor(df$type)

df2 <- df %>% group_by(publication_year, type) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
p <- ggplot(df2, aes(x = publication_year, y = paper, group = type, color = type))
p <- p + geom_point()
p <- p + geom_smooth(method = "loess")
p <- p + ggtitle("Paper consumption per document type")
p <- p + xlab("Year")
p <- p + ylab("Paper consumption")
print(p)

p <- ggplot(df2, aes(x = publication_year, y = n, group = type, color = type))
p <- p + geom_point()
p <- p + geom_smooth(method = "loess")
p <- p + ggtitle("Documents per document type")
p <- p + xlab("Year")
p <- p + ylab("Documents (n)")
print(p)
```


```{r doctypes2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=13, fig.height=6}
df <- mutate(df, length = cut(pagecount, c(0, 32, 72, 148, 250, 450, 800, 2000, 5000, Inf)))

df2 <- df %>% group_by(publication_decade, length) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
p <- ggplot(df2, aes(x = publication_decade, y = paper, group = length, color = length))
p <- p + geom_point()
p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Paper consumption per document type")
p <- p + xlab("Decade")
p <- p + ylab("Paper consumption")
print(p)

p <- ggplot(df2, aes(x = publication_decade, y = n, group = length, color = length))
p <- p + geom_point()
p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Documents per document type")
p <- p + xlab("Decade")
p <- p + ylab("Documents (n)")
print(p)
```


## Top authors

```{r topauth, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=13, fig.height=6}
df <- df.preprocessed
top.authors <- names(rev(sort(table(df.preprocessed$author))))[1:10]
df2 <- df %>% filter(author %in% top.authors) %>% group_by(publication_year, author) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 

p <- ggplot(df2, aes(x = publication_year, y = paper, group = author, color = author))
p <- p + geom_point()
p <- p + geom_line()
#p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Paper consumption per author")
p <- p + xlab("Year")
p <- p + ylab("Paper consumption")
print(p)

p <- ggplot(df2, aes(x = publication_year, y = n, group = author, color = author))
p <- p + geom_point()
p <- p + geom_line()
#p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Documents per author")
p <- p + xlab("Year")
p <- p + ylab("Documents (n)")
print(p)
```




## Nature of the documents over time

Estimated paper consumption by document size

```{r 20150611paris-paper6, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=13, fig.height=7}
df2 <- df %>% group_by(publication_year, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
df2 <- filter(df2, gatherings %in% names(which(table(df2$gatherings) >= 50)))
p <- ggplot(df2, aes(y = paper, x = publication_year, group = gatherings, color = gatherings))
p <- p + geom_point()
p <- p + geom_smooth(method = "loess", size = 1)
p <- p + ggtitle("Annual paper consumption by gatherings")
p <- p + xlab("Year")
p <- p + ylab("Paper consumption")
print(p)
```


Gatherings height: does it change over time (1600's)? How increased printing activity is related to book size trends? Instead, we could use area (height x width), or simply calculate median over time but let's discuss this later.

```{r pagecounts-gatsize, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=7}
df <- df.preprocessed
for (g in c("1to", "2fo", "4to", "8vo")) {
  df2 <- filter(df, publication_year >=1600 & publication_year < 1700 &
       		  !is.na(gatherings.original) &
		  !is.na(height.original) &
		  gatherings.original == g) %>%
       group_by(publication_year, height.original) %>%
       tally()

  p <- ggplot(df2, aes(x = publication_year, y = height.original)) + geom_point(aes(size = n)) 
  p <- p + ggtitle(paste(g, "heights over time"))
  print(p)
}
```


Page counts: does it change over time (1600's)? Also suggested we could calculate some kind of factor for each time period based on this ? In principle, we could calculate this separately for any given publication place as well but leẗ́s discuss this later. Would help to specify some specific places of interest.

```{r pagecounts-gatsize2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=7}
df <- df.preprocessed
for (g in c("1to", "2fo", "4to", "8vo")) {
  df2 <- filter(df, publication_year >=1600 & publication_year < 1700 &
       		  !is.na(gatherings.original) &
		  !is.na(pagecount.orig) &
		  gatherings.original == g)
  df3 <- df2 %>%
       group_by(publication_year, pagecount.orig) %>%
       tally()

  p <- ggplot(df3, aes(x = publication_year, y = pagecount.orig)) + geom_point(aes(size = n))
  p <- p + geom_smooth(data = df2, aes(x = publication_year, y = pagecount.orig))
  p <- p + ggtitle(paste(g, "pagecount over time"))
  print(p)
}
```


Same for documents that have a sufficient number of pages:

```{r pagecounts-gatsize3, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=7}
df <- df.preprocessed
minpages <- 60
for (g in c("1to", "2fo", "4to", "8vo")) {
  df2 <- filter(df, pagecount >= minpages &
       		  !is.na(gatherings.original) &
		  !is.na(pagecount.orig) &
		  gatherings.original == g)
  df3 <- df2 %>%
       group_by(publication_year, pagecount.orig) %>%
       tally()

  p <- ggplot(df3, aes(x = publication_year, y = pagecount.orig)) + geom_point(aes(size = n))
  p <- p + geom_smooth(data = df2, aes(x = publication_year, y = pagecount.orig))
  p <- p + ggtitle(paste(g, "pagecount over time (books with over", minpages, "pages)"))
  print(p)
}
```
